Foundations of ecological and evolutionary change

### 2025-08-04 foundations_of_ecoevo.Rmd
Thanking Andy for comments

Over 120 years ago, @Needham1904 described natural history as "the study of the phenomena of fitness".
Fundamentally, we show why eco-evolutionary change is a statistical interaction between fitness and an individual characteristic.
From this definition, we can recover both population size change and evolutionary change.
Our $\Omega$ can be therefore be interpreted as characteristic-scaled population change, with the special case of $z = 1$ recovering population size change by recognising that each individual contributes one to the total.

@@

This manuscript was supported by joint funding between the French Foundation for Research on Biodiversity (FRB) Centre for the Synthesis and Analysis of Biodiversity (CESAB) and the German Centre for Integrative Biodiversity Research (sDiv).
It was written as part of the Unification of Modern Coexistence Theory and Price Equation (UNICOP) project.
Victor J. Luque was also supported by the Spanish Ministry of Science and Innovation (Project: PID2021-128835NB-I00), and the Conselleria  d’Innovaci$\'{o}$, Universitats, Ci$\`{e}$ncia i Societat Digital – Generalitat  Valenciana (Project: CIGE/2023/16).
We are grateful for many conversations with S$\'{e}$bastien Lion, Kelsey Lyberger, Swati Patel, and especially Lynn Govaert, whose questions helped us clarify the relationship between population growth and fitness.
Brad Duthie would also like to thank Matt Tinsley, Brent Danielson, and Stan Harpole.
Victor J. Luque would also like to thank Lorenzo Baravalle, Pau Carazo, Santiago Ginnobili, Manuel Serra, and Ariel Roff$\'{e}$.
Lastly, we thank two anonymous reviewers.

>>>>

Over 120 years ago, @Needham1904 described natural history as "the study of the phenomena of fitness".
Fundamentally, we show why eco-evolutionary change is a statistical interaction between fitness and individual characteristics.
From this definition, we can recover both population size change and evolutionary change.

@@

This manuscript was supported by joint funding between the French Foundation for Research on Biodiversity (FRB) Centre for the Synthesis and Analysis of Biodiversity (CESAB) and the German Centre for Integrative Biodiversity Research (sDiv).
It was written as part of the Unification of Modern Coexistence Theory and Price Equation (UNICOP) project.
Victor J. Luque was also supported by the Spanish Ministry of Science and Innovation (Project: PID2021-128835NB-I00), and the Conselleria  d’Innovaci$\'{o}$, Universitats, Ci$\`{e}$ncia i Societat Digital – Generalitat  Valenciana (Project: CIGE/2023/16).
We are grateful for many conversations with S$\'{e}$bastien Lion, Kelsey Lyberger, Swati Patel, and especially Lynn Govaert, whose questions helped us clarify the relationship between population growth and fitness.
Brad Duthie would also like to thank Matt Tinsley, Brent Danielson, and Stan Harpole.
Victor J. Luque would also like to thank Lorenzo Baravalle, Pau Carazo, Santiago Ginnobili, Manuel Serra, and Ariel Roff$\'{e}$.
Lastly, we thank two anonymous reviewers and Andy Gardner for helpful comments.
